SBAT: A Simple Browser-based Annotation Tool with Interactive Visualization and Github Support

### 2024-08-31 paper/main.tex
update features

Among the annotation tools, Brat \cite{stenetorp-etal-2012-brat} is a powerful scheme-neutral annotation tool with intuitive and user-friendly interfaces. And it is extremely important for this paper as SBAT is developed upon its UI component. Brat can be used for tasks including but not limited to POS tagging, named entity recognition, semantic role labeling, dependancy and verb frame annotations, and enables user-defined contraints-checking. Besides, it provides a high-quality visualization of annotations by its browser-based UI component implemented in XHTML, Scalable Vector Graphics (SVG) and JavaScript. Last but not the least, Brat integrats a ML-based semantic class disambiguator that outputs multiple annotations with their posibility estimates, which was proven to increase the efficiency of annotators.\\

@@

Another specific-purpose annotation tool is INCEpTION \cite{klie-etal-2018-inception}, which is used for semantic annotation. What differentiates it from the previously-introduced tools is its wide adoption of machine learning models to assist annotation process. It not only utilizes recommenders based on machine learning models to provide users with suggestions for possible labels, but also has an active learning mode where user's feedback can be used to further improve the quality of those suggestions. It also has a comprehensive user management system, where admin can create multiple user accounts for the project and assign them different roles. In terms of the software itself, INCEpTION is a Java application with Spring Boot backend and a web user interface, so its usage requires installation of Java as well as the software.\\

@@

\item Interactive Annotation Interface
The annotation interface of SBAT is completely based on that of BRAT, which is interactive, intuitive and easy-to-use. To add an annotation, one can simply double-click a word or select a span with the mouse, and then select the labels in the pop-up window. The size and display of the interface can be adjusted by the option button, and the labels and their corresponding color can be defined in the JSON config file.
\item web interface, no need to register
SBAT is browser-based, which means one can simply open the website and start annotating in the browser. But unlike Brat, all logics of SBAT are completely executed in the browser, so there is no need to configure a server or install any other software.
\item import and export annotations through ui from github.
Another advantage of SBAT is its support for github. The user only has to enter his own Github Personal Access Token, and can then import files to annotate from his github repositories and export his annotations by making a commit in the same repository. By making use of github commit, the versions of annotations can be well documented and managed. The annotations are exported to a JSON file with attributes like text, tokens, entites.... For file import, all file types are supported, but only annotations in JSON file with these specific attributes can be visualized in the annotation interface, other file types will be simply treated as txt file, i.e. the complete content will be displayed as pure text which can be annotated.
\item potential for collaboration in annotation project
The usage of github repositories to store annotations also enables colloboration and project management. There is a admin config attribute in the JSON config file, which can be modified by the admin to specify the common repository to work on. Then, by sharing the repository to all team members, a team will be able to work on the same annotation project. Different branches can also be created for different members to create and compare multiple versions of annotations.

>>>>

Among the annotation tools, Brat \cite{stenetorp-etal-2012-brat} is a powerful scheme-neutral annotation tool with intuitive and user-friendly interface. And it is extremely important for this paper as SBAT is developed upon its UI component. Brat can be used for tasks including but not limited to POS tagging, named entity recognition, semantic role labeling, dependancy and verb frame annotations, and enables user-defined contraints-checking. Besides, it provides a high-quality visualization of annotations by its browser-based UI component implemented in XHTML, Scalable Vector Graphics (SVG) and JavaScript. Last but not the least, Brat integrats a ML-based semantic class disambiguator that outputs multiple annotations with their posibility estimates, which was proven to increase the efficiency of annotators.\\

@@

Another specialized annotation tool is INCEpTION \cite{klie-etal-2018-inception}, which is used for semantic annotation. What differentiates it from the previously-introduced tools is its wide adoption of machine learning models to assist annotation process. It not only utilizes recommenders based on machine learning models to provide users with suggestions for possible labels, but also has an active learning mode where user's feedback can be used to further improve the quality of those suggestions. It also has a comprehensive user management system, where admin can create multiple user accounts for the project and assign them different roles. In terms of the software itself, INCEpTION is a Java application with Spring Boot backend and a web user interface, so its usage requires installation of Java as well as the software.\\

@@

\subsection{Interactive User Interface}
Based on Brat, the annotation UI of SBAT is responsive, interactive and intuitive to use. To add an annotation, one can simply double-click or select a span with his mouse, and then select corresponding label in the pop-up window. All added labels will be visualized with tags above the annotated text, and different tags differentiate with each other by the user-defined color. And if the user has added comments, they will be shown with a mouse hover. The size and display of the interface can be adjusted by the option button, and the labels and their style are defined in the JSON config file. Besides the main component for annotating text, there is also a step-by-step guide in the beginning for github authentication as well as the export function, all with clear instructions.
\subsection{Browser-Based Application}
SBAT is browser-based, which means one can simply open the website and start annotating in the browser. But unlike Brat, all logics of SBAT are completely executed in the browser, there is no need of software installation or server configuration. One can either store its front-end files in a Github repository to access it via Github pages online, or download the those files to start it locally in the browser. Due to its simple structure, it is fast to run and easy to adapt for different purposes.
\subsection{Github Support}
SBAT can not only be hosted on Github Pages, but also supports import files from or export annotations to Github repositories. To do so, the user only has to enter his Github Personal Access Token (information about the repository name as well as the its owner's user name are entered in advance under "admin config" in the JSON config file), and can then select the github branch as well as the file to open. For importing file, all file formats are supported, and they will be treated as txt file, i.e. the whole content will be displayed as annotation-ready text in the interface. The only exception is JSON file which has been previously exported from SBAT (which has attributes like text, tokens, entities...) -- Their content will be parsed and the annotations saved in the file will be visualized. To export annotations, the user can optionally leave a commit message and click the "commit" button, a JSON file with original texts and annotations will be commited to the same Github repository from which the file was originally imported. By making use of Github commit, the versions of annotations can be well documented and managed.
\subsection{Project Management Support}
SBAT's support for Github also provides possibility for user management in annotation projects. For instance, the admin can create a Github repository for an annotation project and add the repository and owner information in SBAT's config file. By sharing access to this repository with different users, they can contribute to the same project at the same time. Moreover, different branches can be created for different users to create and compare multiple versions. This utilization of Github's file management system facilitates the colloboration of multiple annotators in an annotation project.
